tl22-01 BLASTing the HMMer matches from Plasmodium_Vir (using ORFs from getorf -find 1 option) to Plasmodium vivax P01 v56 proteins

### 2022-10-21 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Add Chapter 5 vivax de novo results folder

chab_tpm.df_pca_coords <- data.frame(sample = rownames(chab_tpm.df_pca_var$coord),
                                        PC1 = chab_tpm.df_pca_coords$Dim.1,
                                        PC3 = chab_tpm.df_pca_coords$Dim.3,
                                        PC2 = chab_tpm.df_pca_coords$Dim.2) %>% 
  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
  mutate( transmission = str_split_fixed(stage, pattern = '_', n = 2)[,1],
          time = str_extract(stage,pattern = '[[:digit:]]{1,2}h|sch.culture'))

colornames <- viridis(length(unique(chab_tpm.df_pca_coords$time)))
colornames <- setNames(colornames, 
                       sort(unique(chab_tpm.df_pca_coords$time)))

@@

```{r}
  cir_plot <- chab_tpm.df %>% 
    filter(Geneid %in% 'PCHAS_0101200') %>% 
    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
    # mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
    mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
           transmission = str_split_fixed(sample, pattern = '_', n = 2)[,1]) %>% 
    group_by(sample, time, transmission) %>% 
    dplyr::summarise(sum_tpm = sum(tpm))

@@

```{r sbp-hi_barchart}
  cir_plot <- chab_tpm.df %>% 
    filter(Geneid %in% "PCHAS_1100300") %>% 
    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
    # mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
    mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
           transmission = str_split_fixed(sample, pattern = '_', n = 2)[,1]) %>% 
    group_by(sample, time, transmission) %>% 
    dplyr::summarise(sum_tpm = sum(tpm))

@@

```{r all_pir_barchart}
#Filter for pirs above 1 TPM in at least one sample
cir_plot <- chab_tpm.df %>% 
  filter(Geneid %in% cir_id,
         if_any(!matches('Geneid'), ~ . > 1)) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
         transmission = str_split_fixed(sample, pattern = '_', n = 2)[,1],
         SorL = cir_clade_info$SorL[match(Geneid, cir_clade_info$Geneid)]) %>% 
  group_by(Geneid, time, transmission, SorL) %>% 
  dplyr::summarise(tpm = median(tpm)) %>% 
  arrange(SorL, tpm)

@@

```{r read_z}
#Can read in the z scores to avoid re-doing all the above.
chab_z.df <- read.csv(file = 'data/Pchabaudi_z.csv')
#Make list object of the TPM values and the z-scores value by concatenating the 
# lists of the zscore dataframes and the TPM lists from above.
chab_z.list <- list('RMT' = c(chab_tpm_avg.list[['RMT']],
                              list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
                                                           pattern = 'Geneid|rMT')])),
                    'SBP' = c(chab_tpm_avg.list[['SBP']],
                              list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
                                                           pattern = 'Geneid|SBP')])))
phaseogram_time_order <- list(rMT = list(zscore_order = list()),
                              SBP = list(zscore_order = list()))

>>>>

chab_tpm.df_pca_coords <- data.frame(sample = rownames(chab_tpm.df_pca_var$coord),
                                        PC1 = chab_tpm.df_pca_coords$Dim.1,
                                        PC3 = chab_tpm.df_pca_coords$Dim.3,
                                        PC2 = chab_tpm.df_pca_coords$Dim.2) %>% 
  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
  mutate( transmission = toupper(str_split_fixed(stage, pattern = '_', n = 2)[,1]),
          time = str_extract(stage,pattern = '[[:digit:]]{1,2}h|sch.culture')) 
# 
# colornames <- viridis(length(unique(chab_tpm.df_pca_coords$time)))
# colornames <- setNames(colornames, 
#                        sort(unique(chab_tpm.df_pca_coords$time)))

@@

```{r}
  cir_plot <- chab_tpm.df %>% 
    filter(Geneid %in% 'PCHAS_0101200') %>% 
    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
    # mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
    mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
           transmission = toupper(str_split_fixed(sample, pattern = '_', n = 2)[,1])) %>% 
    group_by(sample, time, transmission) %>% 
    dplyr::summarise(sum_tpm = sum(tpm))

@@

```{r sbp-hi_barchart}
  cir_plot <- chab_tpm.df %>% 
    filter(Geneid %in% "PCHAS_1100300") %>% 
    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
    # mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
    mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
           transmission = toupper(str_split_fixed(sample, pattern = '_', n = 2)[,1])) %>% 
    group_by(sample, time, transmission) %>% 
    dplyr::summarise(sum_tpm = sum(tpm))

@@

```{r all_pir_barchart}
#Filter for pirs above 1 TPM in at least one sample
cir_plot <- chab_tpm.df %>% 
  filter(Geneid %in% cir_id,
         if_any(!matches('Geneid'), ~ . > 1)) %>% 
  melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
  mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h|sch.culture'),
         transmission = toupper(str_split_fixed(sample, pattern = '_', n = 2)[,1]),
         SorL = cir_clade_info$SorL[match(Geneid, cir_clade_info$Geneid)]) %>% 
  group_by(Geneid, time, transmission, SorL) %>% 
  dplyr::summarise(tpm = median(tpm)) %>% 
  arrange(SorL, tpm)

@@

```{r read_z}
#Can read in the z scores to avoid re-doing all the above.
chab_z.df <- read.csv(file = 'data/Pchabaudi_z.csv')
#Make list object of the TPM values and the z-scores value by concatenating the 
# lists of the zscore dataframes and the TPM lists from above.
# Also rename the RMT dataframe columns to change 'rMT' to 'RMT'
chab_z.list <- list('RMT' = list('experiment' = chab_tpm_avg.list[['RMT']]$experiment,
                                 'tpm' = chab_tpm_avg.list[['RMT']]$tpm %>% 
                                   rename_with(~gsub('rMT', 'RMT', .x)),
                                 'zscores' = chab_z.df[,grepl(colnames(chab_z.df),
                                                              pattern = 'Geneid|rMT')] %>% 
                                   rename_with(~gsub('rMT', 'RMT', .x))),
                    'SBP' = c(chab_tpm_avg.list[['SBP']],
                              list('zscores' = chab_z.df[,grepl(colnames(chab_z.df),
                                                                pattern = 'Geneid|SBP')])))
phaseogram_time_order <- list(rMT = list(zscore_order = list()),
                              SBP = list(zscore_order = list()))

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovoAssemblyAnalysis_SRX697844_SRR1571714_cdhit0.99_blastn.Rmd
Add Chapter 5 vivax de novo results folder

---
title: "tl22-01_VivaxPirsInTheDeNovoAssemblies"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/scspades/SRX697844_SRR1571714/SRX697844_SRR1571714_blastn.outfmt6'
---

@@

```{r low_qcovhsp}
transcripts_covered_by_denovo %>% 
  filter(qcovhsp < 50) %>% 
  kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "300px")
```

>>>>

---
title: "tl22-01_VivaxPirsInTheDeNovoAssemblies SRX697844_SRR1571714_cdhit0.99_blastn.outfmt6"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/scspades/SRX697844_SRR1571714/SRX697844_SRR1571714_cdhit0.99_blastn.outfmt6'
---

@@

```{r low_qcovhsp}
if(transcripts_covered_by_denovo %>% 
   filter(qcovhsp < 50) %>% 
   nrow() > 0){
  transcripts_covered_by_denovo %>% 
    filter(qcovhsp < 50) %>% 
    kbl() %>% kable_styling() %>% scroll_box(width = "100%", height = "300px")
}
```

# All the BLAST matches

```{r transcripts_coverage_allHits_transcripts_graph}
blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'aa', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

# BLAST hit coverages for the top alignments.

```{r transcripts_coverage_topBLAST_transcripts_graph}
blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  group_by(sseqid) %>% 
  slice_min(evalue, n = 1) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'aa', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovoAssemblyAnalysis_SRX697844_SRR1571714_scspades_blastpP01-Zhu16-Sal1vir.Rmd
Add Chapter 5 vivax de novo results folder

---
title: "tl22-01_VivaxPirsInTheDeNovoAssemblies SRX697844_SRR1571714_cdhit0.99_blastn.outfmt6"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/scspades/SRX697844_SRR1571714/SRX697844_SRR1571714_cdhit0.99_blastn.outfmt6'
---

@@

This script is looks at the BLAST results of a Trinity _de novo_ assembled transcriptome.

@@

*The level of counts / TPM of pirs that are detected by the Trinity de novo assembled transcriptome.
**e.g. What is the minimum expressed gene detected / What is the maximally expressed gene not detected.

# To-do:
*Check how they BLAST to the rest of the transcriptome, do pir de novo transcripts also BLAST to non-pir transcripts?
*Select the most informative data and poo-poo the rest.
*Save useful files made by this script as csv.

@@

blast <- read.table(params$blast_output_file,
                    col.names = outfmt6_headers) %>% 
  mutate(
         sseqid = unlist(
           lapply(
             str_extract_all(sseqid, 
                             pattern = '.+(?=\\.1)'), 
             paste, 
             collapse = '_')
         ),
         scovhsp = round( (nident/slen * 100), digits = 2),
         scovs = round(((abs(send - sstart) + 1)/slen *100), digits = 2),
         evalue = format(evalue, scientific = TRUE)
  ) %>% 
  #Filter for the pir genes.
  filter(nident > filter_nident,
         sseqid %in% pir.info$Gene.ID) %>%
  group_by(qseqid) %>% 
  mutate(num_genes_match = n_distinct(sseqid))

@@

@@

There are `r length(unique(transcripts_covered_by_denovo$sseqid))` known transcripts covered by >=95% of at least one individual _de novo_ transcript. There is a mean of `r mean(transcripts_covered_by_denovo$mismatch)` mismatches and `r mean(transcripts_covered_by_denovo$gapopen)` gaps and a median of `r median(transcripts_covered_by_denovo$mismatch)` mismatches and `r median(transcripts_covered_by_denovo$gapopen)` gaps in these alignments (Note that there are probably some de novo transcripts that align >=95% to the same known transcript). Of these `r length(unique(transcripts_covered_by_denovo$sseqid[transcripts_covered_by_denovo$num_match_all == 1]))`
 match only that known transcript and no others by the BLAST search. `r length(unique(filter(ungroup(transcripts_covered_by_denovo), qcovhsp < 50)$sseqid))` of the genes with >=95% known transcript alignments cover less than 50% of the de novo transcript.

>>>>

---
title: "tl22-01_VivaxPirsInTheDeNovoAssemblies SRX697844_SRR1571714_getorfFind1_blastp-P01-Zhu16_Sal1vir"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/blast_to_Proteins/scspades_blastpP01-Zhu16-Sal1vir/SRX697844_SRR1571714/SRX697844_SRR1571714_scspades_blastpP01-Zhu16-Sal1vir_hmmerPfam-Plasmodium_Vir_blastp-P01-all_Sal1-vir_Zhu2016Vir.outfmt6'
---

@@

This script is looks at the BLAST results of the HMMer matches with the Plasmodium_Vir among the ORFs from the Plasmodium vivax RNAseq scspades assembly. BLASTed to the Plasmodium vivax P01 PlasmoDB v56 proteins, the Zhu 2016 de novo mRNA vir predicted getorfs, and the Sal1 virs.

@@

# To-do:

@@

blast <- read.table(params$blast_output_file,
                    col.names = outfmt6_headers) %>% 
  mutate(

         scovhsp = round( (nident/slen * 100), digits = 2),
         scovs = round(((abs(send - sstart) + 1)/slen *100), digits = 2),
         evalue = format(evalue, scientific = TRUE)
  ) %>% 
  #Filter for the pir genes.
  # The Sal1 and Zhu genes are already only the virs so I grep for them first in ifelse.
  # Then, for the P01, I extract the gene name and search in the pir.info list.
  filter(nident > filter_nident,
         ifelse(grepl(sseqid, pattern = 'PVX|comp'),
                TRUE,
                str_extract_all(sseqid, 
                                pattern = '.+(?=\\.1)') %in% pir.info$Gene.ID)) %>%
  group_by(qseqid) %>% 
  mutate(num_genes_match = n_distinct(sseqid))

@@

# BLAST hit coverages for the top alignments to the Zhu 2016 de novo virs.

```{r transcripts_coverage_topBLAST_Zhu16transcripts_graph}
blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  filter(grepl(sseqid, 
               pattern = 'comp')) %>% 
  group_by(sseqid) %>% 
  slice_min(evalue, n = 1) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'aa', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

@@

There are `r length(unique(transcripts_covered_by_denovo$sseqid))` known transcripts covered by >=95% of at least one individual _de novo_ transcript. Of these `r length(unique(str_subset(transcripts_covered_by_denovo$sseqid, pattern = 'comp')))` are predicted virs from Zhu et al, 2016 and `r length(unique(str_subset(transcripts_covered_by_denovo$sseqid, pattern = 'PVX')))` are from Sal1 genome. There is a mean of `r mean(transcripts_covered_by_denovo$mismatch)` mismatches and `r mean(transcripts_covered_by_denovo$gapopen)` gaps and a median of `r median(transcripts_covered_by_denovo$mismatch)` mismatches and `r median(transcripts_covered_by_denovo$gapopen)` gaps in these alignments (Note that there are probably some de novo transcripts that align >=95% to the same known transcript). Of these `r length(unique(transcripts_covered_by_denovo$sseqid[transcripts_covered_by_denovo$num_match_all == 1]))`
 match only that known transcript and no others by the BLAST search. `r length(unique(filter(ungroup(transcripts_covered_by_denovo), qcovhsp < 50)$sseqid))` of the genes with >=95% known transcript alignments cover less than 50% of the de novo transcript.

### 2022-10-21 chapter5_results_vivaxDeNovo/tl22-01_VvxDeNovo_getorfFind1_hmmerMatches_blast2VvxProt.Rmd
Add Chapter 5 vivax de novo results folder

---
title: "tl22-01_VivaxPirsInTheDeNovoAssemblies SRX697844_SRR1571714_getorfFind1_blastp-P01-Zhu16_Sal1vir"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/blast_to_Proteins/scspades_blastpP01-Zhu16-Sal1vir/SRX697844_SRR1571714/SRX697844_SRR1571714_scspades_blastpP01-Zhu16-Sal1vir_hmmerPfam-Plasmodium_Vir_blastp-P01-all_Sal1-vir_Zhu2016Vir.outfmt6'
---

@@

This script is looks at the BLAST results of the HMMer matches with the Plasmodium_Vir among the ORFs from the Plasmodium vivax RNAseq scspades assembly. BLASTed to the Plasmodium vivax P01 PlasmoDB v56 proteins, the Zhu 2016 de novo mRNA vir predicted getorfs, and the Sal1 virs.

# What this report shows:

# To-do:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)

#Possible files for blast_output_file:
# '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/scspades/SRX697844_SRR1571714/SRX697844_SRR1571714_blastn.outfmt6'
```

@@

```{r read_blast_results}
# list.files('../blast', pattern = 'outfmt')

#Set a filter for bp length to assume is a valid match
filter_nident <- 100

@@

blast <- read.table(params$blast_output_file,
                    col.names = outfmt6_headers) %>% 
  mutate(

         scovhsp = round( (nident/slen * 100), digits = 2),
         scovs = round(((abs(send - sstart) + 1)/slen *100), digits = 2),
         evalue = format(evalue, scientific = TRUE)
  ) %>% 
  #Filter for the pir genes.
  # The Sal1 and Zhu genes are already only the virs so I grep for them first in ifelse.
  # Then, for the P01, I extract the gene name and search in the pir.info list.
  filter(nident > filter_nident,
         ifelse(grepl(sseqid, pattern = 'PVX|comp'),
                TRUE,
                str_extract_all(sseqid, 
                                pattern = '.+(?=\\.1)') %in% pir.info$Gene.ID)) %>%
  group_by(qseqid) %>% 
  mutate(num_genes_match = n_distinct(sseqid))

@@

# Coverage of the subject (known transcripts) and query (assembled de novo transcripts) sequences.

Using only the de novo transcripts that uniquely match one known pir transcript.

For our purposes the subject coverage is probably most important.

```{r scov_v_qcov}
fig <- plot_ly(
  blast_uniq,
  x = ~qcovhsp,
  y = ~scovhsp,
  text = ~paste(sseqid)
) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
fig
```

Number of unique genes found by BLAST: `r unique(blast$sseqid) %>% length`.

@@

```{r transcripts_covering_most_of_known_transcripts_graph}
blast %>% 
  filter(qseqid %in% unique(transcripts_covered_by_denovo$qseqid)) %>%
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'bp', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

@@

# BLAST hit coverages for the top alignments to the Zhu 2016 de novo virs.

```{r transcripts_coverage_topBLAST_Zhu16transcripts_graph}
blast %>% 
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  filter(grepl(sseqid, 
               pattern = 'comp')) %>% 
  group_by(sseqid) %>% 
  slice_min(evalue, n = 1) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'aa', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

@@

transcripts_covered_by_denovo_diffcov <- lapply(seq(0,100,by = 5),
                                                function(covrg) {
                                                  blast %>% 
                                                    group_by(sseqid) %>% 
                                                    filter(scovhsp >= covrg) %>%
                                                    .$sseqid %>% unique %>% length
                                                })
names(transcripts_covered_by_denovo_diffcov) <- seq(0,100,by = 5)
plot(y = unlist(transcripts_covered_by_denovo_diffcov), 
     x = seq(0,100,by = 5),
     ylim = c(0,nrow(pir.info)),
     main = 'For all de novo transcripts',
     ylab = 'Number of known transcripts covered',
     xlab = 'Percentage of coverage by individual transcripts',
     yaxt = 'none')
axis(2, at = seq(0,nrow(pir.info),by = 100))
# abline(a = 25, b = 0)

@@

There are `r length(unique(transcripts_covered_by_denovo$sseqid))` known transcripts covered by >=95% of at least one individual _de novo_ transcript. Of these `r length(unique(str_subset(transcripts_covered_by_denovo$sseqid, pattern = 'comp')))` are predicted virs from Zhu et al, 2016 and `r length(unique(str_subset(transcripts_covered_by_denovo$sseqid, pattern = 'PVX')))` are from Sal1 genome. There is a mean of `r mean(transcripts_covered_by_denovo$mismatch)` mismatches and `r mean(transcripts_covered_by_denovo$gapopen)` gaps and a median of `r median(transcripts_covered_by_denovo$mismatch)` mismatches and `r median(transcripts_covered_by_denovo$gapopen)` gaps in these alignments (Note that there are probably some de novo transcripts that align >=95% to the same known transcript). Of these `r length(unique(transcripts_covered_by_denovo$sseqid[transcripts_covered_by_denovo$num_match_all == 1]))`
 match only that known transcript and no others by the BLAST search. `r length(unique(filter(ungroup(transcripts_covered_by_denovo), qcovhsp < 50)$sseqid))` of the genes with >=95% known transcript alignments cover less than 50% of the de novo transcript.

>>>>

---
title: "tl22-01 BLASTing the HMMer matches from Plasmodium_Vir (using ORFs from getorf -find 1 option) to Plasmodium vivax P01 v56 proteins"
output: 
  html_document:
    theme: cosmo
    code_download: true
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  blast_output_file: '/camp/lab/langhornej/working/HPC/littlet/tl22-01_vivaxDeNovo/blast/blast_to_Proteins/scspades_getorfTableZero/SRX697844_SRR1571714/SRX697844_SRR1571714_scspades_getorfTableZero_hmmerPfam-Plasmodium_Vir_bast-VvxProt.outfmt6'
---

@@

This script is looks at the BLAST results of the HMMer matches with the Plasmodium_Vir among the ORFs from the Plasmodium vivax RNAseq scspades assembly. BLASTed to the Plasmodium vivax P01 PlasmoDB v56 proteins.

# To-do:
* BLASTs to non-pirs?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

@@

```{r read_blast_results}
#Set a filter for bp length to assume is a valid match
filter_nident <- 100

@@

blast_all <- read.table(params$blast_output_file,
                        col.names = outfmt6_headers) %>% 
  mutate(
    sseqid = unlist(
      lapply(
        str_extract_all(sseqid, 
                        pattern = '.+(?=\\.1)'), 
        paste, 
        collapse = '_')
    ),
    scovhsp = round( (nident/slen * 100), digits = 2),
    scovs = round(((abs(send - sstart) + 1)/slen *100), digits = 2),
    evalue = format(evalue, scientific = TRUE)
  ) %>% 
  filter(nident > filter_nident) %>%
  group_by(qseqid) %>% 
  mutate(num_genes_match = n_distinct(sseqid))

#Filter for the pir genes.
blast <- filter(blast_all, 
                sseqid %in% pir.info$Gene.ID) %>%
  group_by(qseqid) %>% 
  mutate(num_genes_match = n_distinct(sseqid))

#Non-pir genes
blast_nonpir <- filter(blast_all, 
                       !sseqid %in% pir.info$Gene.ID) %>%
  group_by(qseqid) %>% 
  mutate(num_genes_match = n_distinct(sseqid))

@@

@@

Note: PVP01_1147700 = ancestral

```{r transcripts_covering_most_of_known_transcripts_graph}
blast %>% 
  filter(qseqid %in% unique(transcripts_covered_by_denovo$qseqid)) %>%
  group_by(qseqid) %>%
  mutate(num_match_all = n_distinct(sseqid)) %>% 
  mutate(qseqid = paste0(qseqid, '; n = ', num_match_all),
         over95 = ifelse(scovhsp >= 95, 
                         '>=95% coverage', 
                         '<95% coverage')) %>% 
  plot_ly(x = ~qcovhsp,
          y = ~scovhsp,
          text = ~paste(sseqid, '\n', 
                        length, 'aa', '\n',
                        scovhsp, '% coverage of known'),
          color = ~qseqid,
          size = ~length,
          symbol = ~over95,
          symbols = c('triangle-up', 'circle'),
          sizes = c(2,100),
          legendgroup = ~qseqid) %>% 
  layout(xaxis = list(title = "De novo transcript coverage"),
         yaxis = list(title = "Known transcript coverage"))
```

@@

@@

transcripts_covered_by_denovo_diffcov <- lapply(seq(0,100,by = 5),
                                                function(covrg) {
                                                  blast %>% 
                                                    group_by(sseqid) %>% 
                                                    filter(scovhsp >= covrg) %>%
                                                    .$sseqid %>% unique %>% length
                                                })
names(transcripts_covered_by_denovo_diffcov) <- seq(0,100,by = 5)
plot(y = unlist(transcripts_covered_by_denovo_diffcov), 
     x = seq(0,100,by = 5),
     ylim = c(0,nrow(pir.info)),
     main = 'For pir de novo transcripts',
     ylab = 'Number of known transcripts covered',
     xlab = 'Percentage of coverage by individual transcripts',
     yaxt = 'none')
axis(2, at = seq(0,nrow(pir.info),by = 100))
# abline(a = 25, b = 0)

@@

There are `r length(unique(transcripts_covered_by_denovo$sseqid))` known transcripts covered by >=95% of at least one individual _de novo_ transcript. There is a mean of `r mean(transcripts_covered_by_denovo$mismatch)` mismatches and `r mean(transcripts_covered_by_denovo$gapopen)` gaps and a median of `r median(transcripts_covered_by_denovo$mismatch)` mismatches and `r median(transcripts_covered_by_denovo$gapopen)` gaps in these alignments (Note that there are probably some de novo transcripts that align >=95% to the same known transcript). Of these `r length(unique(transcripts_covered_by_denovo$sseqid[transcripts_covered_by_denovo$num_match_all == 1]))`
 match only that known transcript and no others by the BLAST search. `r length(unique(filter(ungroup(transcripts_covered_by_denovo), qcovhsp < 50)$sseqid))` of the genes with >=95% known transcript alignments cover less than 50% of the de novo transcript.
